---
title: "Getting and Cleaning data_project"

output: html_document
---

####Feb 2016

### Instruction

##### The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis. You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 1) a tidy data set as described below, 2) a link to a Github repository with your script for performing the analysis, and 3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called CodeBook.md. You should also include a README.md in the repo with your scripts. This repo explains how all of the scripts work and how they are connected.
##### One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithmsto attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone.  




```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}

library(plyr)
library(dplyr)
library(data.table)
library(tidyr)
library("memisc", lib.loc="~/R/win-library/3.2")
##Download the file 

    if(!file.exists("./data")){dir.create("./data")}
        fileurl<-("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip")
        download.file(url=fileurl,destfile="C:/Users/afarn_000.AZADEHSURFACE/Documents/R/data/Dataset.zip",mode="w")

## Unzip the file 
   file <- unzip(zipfile="C:/Users/afarn_000.AZADEHSURFACE/Downloads/Compressed/getdata_projectfiles_UCI HAR Dataset.zip",exdir = "./data")

path<-file.path("./data","UCI HAR Dataset")

lists<-list.files(path,recursive = TRUE)
lists
```

###Read data file and creat a data

```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}
#Read Activity file from train and test folder
Activitytrain<- tbl_df(read.table(file.path(path,"train","Y_train.txt"), header = FALSE))
Activitytest<- tbl_df(read.table(file.path(path,"test","Y_test.txt"), header = FALSE))

#Read features file from train and test folder
Featuretrain<- tbl_df(read.table(file.path(path,"train","X_train.txt"), header = FALSE))
Featuretest<- tbl_df(read.table(file.path(path,"test","X_test.txt"), header = FALSE))


#Read Subject file from train and test folder
Subjecttrain<- tbl_df(read.table(file.path(path,"train","subject_train.txt"), header = FALSE))
Subjecttest<- tbl_df(read.table(file.path(path,"test","subject_test.txt"), header = FALSE))

```

### Merges the training and the test sets to create one data set

```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}

#combine by row
Activitydata<-rbind(Activitytrain,Activitytest)
Subjectdata<-rbind(Subjecttrain,Subjecttest)
Featuredata<-rbind(Featuretrain,Featuretest)


#Setnames for Feature ,Subject and Activity table
Featurenames <- tbl_df(read.table(file.path(path, "features.txt"),header = FALSE))

names(Featuredata)<-Featurenames$V2
names(Activitydata)<-c("Activity")
names(Subjectdata)<-c("Subject")


#combine by column
temp<-cbind(Subjectdata,Activitydata)
Database<-cbind(temp,Featuredata)

str(Database)   
```

### Extracts only the measurements on the mean and standard deviation for each measurement

```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}

# Subset measurements for the mean and standard deviation.
 
FeaturesNamessubset<-Featurenames$V2[grep("mean\\(\\)|std\\(\\)", Featurenames$V2)]

#Subset the data frame by selected names of Features

selected_names<-c(as.character(FeaturesNamessubset), "Subject", "Activity" )
Database<-subset(Database,select=selected_names)

names(Database)

```

###Uses descriptive activity names to name the activities in the data set

```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}
#Descriptive activity for Activity_label
Activity_labels <- tbl_df(read.table(file.path(path, "activity_labels.txt"),header = FALSE))

Database$Activity<- factor(Database$Activity,labels=as.character(Activity_labels$V2))

names(Database)
```

###Appropriately labels the data set with descriptive variable names

```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}

# prefix t is replaced by time
# prefix f is replaced by frequency
# Acc is replaced by Accelerometer measurement
# Gyro is replaced by Gyroscope measurement
# Mag is replaced by Magnitude of movement
# Bodyis replaced by Body movement

names(Database)<-gsub("^t", "time", names(Database))
names(Database)<-gsub("^f", "frequency", names(Database))
names(Database)<-gsub("Acc", "Accelerometer", names(Database))
names(Database)<-gsub("Gyro", "Gyroscope", names(Database))
names(Database)<-gsub("Mag", "Magnitude", names(Database))
names(Database)<-gsub("BodyBody", "Body", names(Database))

names(Database)
```

###From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r,echo=TRUE, results='asis', warning=TRUE, message=TRUE}


New_Database<-aggregate(. ~Subject + Activity, data=Database, mean)

New_Database<-New_Database[order(New_Database$Subject,New_Database$Activity),]
write.table(New_Database, file = "tidydata.txt",row.name=FALSE)

codebook(Database)
Write(codebook(Database), file="Data-cdbk.txt")

```


